---
title: "Cognitive Differences in Reward Processing between ADHD and Neurotypical Children"
author: "Vidushi Kataria"
date: "`r Sys.Date()`"
output:
  ioslides_presentation:
    theme: cerulean
---

```{r load, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tidyverse)
library(readr)
```

## Research Objective

- How does reward processing differ between children diagnosed with ADHD and neurotypical children?

## Fork Question

- How does the timing and accuracy of responses during reward trials in the n-back task differ between children with ADHD and neurotypical children? 


## Data Source 
- Public datasets accessed from OpenNeuro.org: https://openneuro.org/datasets/ds002424/versions/1.2.0 


## Method
- Participants: Children with ADHD (5) and neurotypical children (5)
- Experiment: Participants performed an n-back task, identifying whether a letter matched either by identity (verbal) or location (spatial)
    - Conditions: Small vs. large rewards and immediate vs. delayed feedback
    - Responses: Accuracy and timing of button-box presses analyzed during reward trials to compare groups
- Design: Block design alternating between experimental (1-back, 2-back) and fixation blocks.


```{r events01_files, include=FALSE}
adhd_subjects <- c("sub-01", "sub-02", "sub-04", "sub-05", "sub-17")
neurotypical_subjects <- c("sub-03", "sub-18", "sub-19", "sub-20", "sub-21")

sub01_dir <- "/Users/vidushikataria/Downloads/CogSci412_Final_VK/final_proj/sub-01"

events01_files <- list.files(
  path = sub01_dir, 
  pattern = "_events.tsv$", 
  full.names = TRUE
)

sub01_data <- lapply(events01_files, function(file) {
  df <- read_tsv(file)
  task_name <- sub(".*task-(.*)_events.tsv$", "\\1", basename(file))
  df$task <- task_name
  
  return(df)
}) %>%
  bind_rows()  
```


```{r events02_files, include=FALSE}
sub02_dir <- "/Users/vidushikataria/Downloads/CogSci412_Final_VK/final_proj/sub-02"

events02_files <- list.files(
  path = sub02_dir, 
  pattern = "_events.tsv$", 
  full.names = TRUE
)

sub02_data <- lapply(events02_files, function(file) {
  df <- read_tsv(file)
  task_name <- sub(".*task-(.*)_events.tsv$", "\\1", basename(file))
  df$task <- task_name
  
  return(df)
}) %>%
  bind_rows()  
```


```{r events03_files, include=FALSE}
sub03_dir <- "/Users/vidushikataria/Downloads/CogSci412_Final_VK/final_proj/sub-03"

events03_files <- list.files(
  path = sub03_dir, 
  pattern = "_events.tsv$", 
  full.names = TRUE
)

sub03_data <- lapply(events03_files, function(file) {
  df <- read_tsv(file)
  task_name <- sub(".*task-(.*)_events.tsv$", "\\1", basename(file))
  df$task <- task_name
  
  return(df)
}) %>%
  bind_rows()  
```


```{r events04_files, include=FALSE}
sub04_dir <- "/Users/vidushikataria/Downloads/CogSci412_Final_VK/final_proj/sub-04"

events04_files <- list.files(
  path = sub04_dir, 
  pattern = "_events.tsv$", 
  full.names = TRUE
)

sub04_data <- lapply(events04_files, function(file) {
  df <- read_tsv(file)
  task_name <- sub(".*task-(.*)_events.tsv$", "\\1", basename(file))
  df$task <- task_name
  
  return(df)
}) %>%
  bind_rows()  
```


```{r events05_files, include=FALSE}
sub05_dir <- "/Users/vidushikataria/Downloads/CogSci412_Final_VK/final_proj/sub-05"

events05_files <- list.files(
  path = sub05_dir, 
  pattern = "_events.tsv$", 
  full.names = TRUE
)

sub05_data <- lapply(events05_files, function(file) {
  df <- read_tsv(file)
  task_name <- sub(".*task-(.*)_events.tsv$", "\\1", basename(file))
  df$task <- task_name
  
  return(df)
}) %>%
  bind_rows()  
```

`
```{r events17_files, include=FALSE}
sub17_dir <- "/Users/vidushikataria/Downloads/CogSci412_Final_VK/final_proj/sub-17"

events17_files <- list.files(
  path = sub17_dir, 
  pattern = "_events.tsv$", 
  full.names = TRUE
)

sub17_data <- lapply(events17_files, function(file) {
  df <- read_tsv(file)
  task_name <- sub(".*task-(.*)_events.tsv$", "\\1", basename(file))
  df$task <- task_name
  
  return(df)
}) %>%
  bind_rows()  
```


```{r events18_files, include=FALSE}
sub18_dir <- "/Users/vidushikataria/Downloads/CogSci412_Final_VK/final_proj/sub-18"

events18_files <- list.files(
  path = sub18_dir, 
  pattern = "_events.tsv$", 
  full.names = TRUE
)

sub18_data <- lapply(events18_files, function(file) {
  df <- read_tsv(file)
  task_name <- sub(".*task-(.*)_events.tsv$", "\\1", basename(file))
  df$task <- task_name
  
  return(df)
}) %>%
  bind_rows()  
```


```{r events19_files, include=FALSE}
sub19_dir <- "/Users/vidushikataria/Downloads/CogSci412_Final_VK/final_proj/sub-19"

events19_files <- list.files(
  path = sub19_dir, 
  pattern = "_events.tsv$", 
  full.names = TRUE
)

sub19_data <- lapply(events19_files, function(file) {
  df <- read_tsv(file)
  task_name <- sub(".*task-(.*)_events.tsv$", "\\1", basename(file))
  df$task <- task_name
  
  return(df)
}) %>%
  bind_rows()  
```


```{r events20_files, include=FALSE}
sub20_dir <- "/Users/vidushikataria/Downloads/CogSci412_Final_VK/final_proj/sub-20"

events20_files <- list.files(
  path = sub20_dir, 
  pattern = "_events.tsv$", 
  full.names = TRUE
)

sub20_data <- lapply(events20_files, function(file) {
  df <- read_tsv(file)
  task_name <- sub(".*task-(.*)_events.tsv$", "\\1", basename(file))
  df$task <- task_name
  
  return(df)
}) %>%
  bind_rows()  
```


```{r events21_files, include=FALSE}
sub21_dir <- "/Users/vidushikataria/Downloads/CogSci412_Final_VK/final_proj/sub-21"

events21_files <- list.files(
  path = sub21_dir, 
  pattern = "_events.tsv$", 
  full.names = TRUE
)

sub21_data <- lapply(events21_files, function(file) {
  df <- read_tsv(file)
  task_name <- sub(".*task-(.*)_events.tsv$", "\\1", basename(file))
  df$task <- task_name
  
  return(df)
}) %>%
  bind_rows()  
```


```{r wrangling, include=FALSE}
adhd_data <- bind_rows(sub01_data, sub02_data, sub04_data, sub05_data, sub17_data) %>%
  mutate(group = "ADHD")

neurotypical_data <- bind_rows(sub03_data, sub18_data, sub19_data, sub20_data, sub21_data) %>%
  mutate(group = "Neurotypical")

adhd_data$group <- "ADHD"
neurotypical_data$group <- "Neurotypical"

combined_data <- bind_rows(adhd_data, neurotypical_data)

combined_data <- combined_data %>%
  mutate(response_time = as.numeric(response_time))
sum(is.na(combined_data$response_time))
```


```{r filter-reward, include=FALSE}
reward_data <- combined_data %>%
  filter(stimulus == "redcross.bmp") %>%  
  group_by(group, task, trial_type) %>% 
  summarize(
    mean_acc = mean(accuracy, na.rm = TRUE),
    mean_rt = mean(response_time, na.rm = TRUE),
    .groups = "drop"
  ) %>%
  rename(reward_amount = task, feedback_type = trial_type)
```



```{r plot-1, include=FALSE}
no1_plot <- ggplot(reward_data, aes(x = reward_amount, y = mean_acc, fill = group)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  facet_wrap(~feedback_type) +
  labs(
    title = "Accuracy by Reward Amount and Feedback Type",
    x = "Reward Amount",
    y = "Mean Accuracy",
    fill = "Group"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(size = 12, face = "bold")
  )
```


## Figure 1
```{r fig-1, include=TRUE}
print(no1_plot)
```
Neurotypical group generally has higher accuracy than the ADHD group, especially at higher reward amounts and for different feedback types (baseline, VLD, VSD). 


```{r plot-2, include=FALSE}
no2_plot <- ggplot(reward_data, aes(x = reward_amount, y = mean_rt, fill = group)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  facet_wrap(~feedback_type) +
  labs(
    title = "Response Time by Reward Amount and Feedback Type",
    x = "Reward Amount",
    y = "Mean Response Time (s)",
    fill = "Group"
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    strip.text = element_text(size = 12, face = "bold")
  )
```


## Figure 2
```{r fig-2, include=TRUE}
print(no2_plot)
```
ADHD group tends to have longer response times compared to the neurotypical group, particularly at higher reward amounts.


## What does this mean?
- ADHD and neurotypical groups exhibit distinct patterns in terms of both the accuracy and timing of their responses during the n-back task, especially when the task involves varying reward amounts and feedback.
- Could imply differences in cognitive processing speed and decision-making when rewards are involved.




